3D Object Classification Based on RGBD Images

### 2014-10-12 docs/min/kkr13-eng.tex
min

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% This file is delivered as an usage example of the file 'kkr13-eng.sty'.
% The style file 'kkr13-eng.sty' has been prepared for papers intended 
% to be presented at the XIII National Conference on Robotics in Poland.
%
% The file 'kkr13-eng.sty' contains definitions of macros for English
% and Polish authors. Thus their names are in English and Polish respectively.
% They define the same operations. English and Polish names are used
% only for authors' convenience. 
% Therefore both types of macros can be used in this type of the document.
% The same type of macros are defined in the style file defined for papers 
% written in Polish.
% Below there is a list of macros names in English and their equivalent forms
% in Polish.
%
%       \articleTitle          =   \tytulArtykulu
%       \titleFootNote         =   \stopkaPrzypisTytulu
%       \articleAuthor         =   \autorArtykulu
%       \affiliation           =   \instytucja
%       \theSameAffiliationAs  =   \instytucjaTaSamaJak
%       \articleShortTitle     =   \naglowekTytulSkrocony
%       \breakAuthorsLine      =   \zlamLinieAutorow
%       \authorsForHeader      =   \naglowekAutorzyArtykulu
%       \authorsCustomizedList =   \wlasnaListaAutorow
%       \abstract              =   \streszczenie
%       \Equation              =   \wzor
%       \refeq                 =   \refwzor
%       \reffig                =   \refrys
%       \refFig                =   \refRys
%
% Author: Bogdan Kreczmer 
% Last modification:      2013.12.09
% 
% In the case of any doubt, please, contact via e-mail.
% The address of the author is:
%  bogdan.kreczmer@pwr.wroc.pl
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt,twoside]{article}
\usepackage{kkr13-eng}
\usepackage{todonotes}
\usepackage[hidelinks]{hyperref}
\usepackage{array}
\usepackage{booktabs}
\usepackage{cite}
\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subscript}[1]{\ensuremath{_{\textrm{#1}}}}

\articleTitle{3D Object Classification \\ Based on RGBD Images}

%
% For information of a project (if any)
%
\titleFootNote{The work described in this paper was conducted within author's Bachelor's dissertation}

%
% Parameters:
%   1 - initials of the first and middle name of an author
%   2 - the first and middle name of an author
%   3 - the last name of an author
%
\articleAuthor{A.}{Adam}{Kosiorek}

\affiliation[institute1]{ Institute of Automation and Robotics, The Faculty of Mechatronics of Warsaw University of\\ Technology, email: iair@mchtr.pw.edu.pl, website: http://iair.mchtr.pw.edu.pl}

@@

%\listoftodos

@@

  This paper introduces a Bag of Words (BoW) semantic object classification framework based on RGBD data registered by a Microsoft Kinect camera. The BoW concept in natural language processing is introduced and a BoW image processing pipeline is described. Each part of the pipeline (region detection, feature extraction and vector quantization) is characterized and the most successful algorithms are outlined. Applications of BoW in image processing and their results are discussed. Additionally, the problem of classification is described. Algorithms with high performance potential were identified and evaluated. Salient region detection is done by SIFT, ISS3D and Uniform Sampling. Description is performed by FPFH, PFH and PFHRGB. Vector quantization was carried out by the KMeans algorithm. An SVM with the RBF kernel trained with an All-vs-All was used as a classifier. We achieved the accuracy of 65.22\% on 8 categories of the Berkely 3D Object Dataset and 62.3\% on a 10 category dataset compiled by Zhang from the Tokyo University.
}

\maketitle

\section{ INTRODUCTION }

  It has never been easier to capture visual information. Popular storage services grow lighting fast due to terabytes of photographs and movies uploaded every day. The growth is so fast that hand-tagging and description, the traditional means of annotation, cease to suffice. They are ambiguous, emotional and rarely optimal. With no better solution at hand databases are becoming increasingly harder to browse. Another, but quite similar setting occurs in the field of mobile robotics. A mobile robot is there to interact with its environment. If so, it would be desirable for the robot to know what kind of surroundings it is in. The problem can be treated as a scene or object classification problem, one of the most popular computer vision issues in the last decade \cite{ponce2011cv}.

  This paper tackles two issues. Firstly, the Bag of Words (or BoW for short) approach to image classification is reviewed. The Bag of Words method originated from the natural language processing domain \cite{tsai2012bag}. It makes a strong assumption that word occurrence in a text document define its meaning. It were Csurka \textit{et al} who first used BoW for image classification in 2004 \cite{csurka2004visual}. Since images cannot be simply cast into natural language documents a pipeline that help us achieve this goal is needed. It usually consist of: keypoint detection, feature extraction, vector quantization together with codebook construction and classification, steps that will be discussed later.

  Being efficient, effective and flexible, a Bag of Words processing pipeline is, however, a blessing in a disguise. While in fact it can provide satisfactory results with little effort, it is cumbersome to tune, thus spectacular performance is hard to achieve. One reason is the abundance of suitable algorithms available. Wide variety of building blocks might make it flexible, but it renders choosing an optimal combination infeasible.

  The rest of this work is organised as follows: Section 2 describes the Bag of Words technique in detail. Construction of an image model is discussed and the most popular algorithms are outlined. Then, applications of BoW in computer vision are reported and the problem of classification is introduced. Section 3 describes two renown RGBD image datasets suitable for evaluation of RGBD data-based object classifiers. Section 4 addresses the topics of experimental setup, conducted experiments and provides the results. Finally, Section 5 concludes this work.

\section{ BAG OF WORDS }

  The Bag of Words model is an intermediate representation in a form of a histogram used in natural language processing, information retrieval and data mining. It obliterates any grammatical information as well as word order so as to preserve word incidence only.

@@

  \refFig{fig:bow_example} shows a piece of text and its BoW model, with articles and punctuation marks left out. In order to compare different documents a global dictionary (codebook) must be built. It is constructed by taking every unique word from all available documents. One can image that if a dictionary is of any considerable size the resulting representation of especially small documents can be very sparse. Classification algorithms can be optimised for sparse data in order to boost performance.

  In natural language processing BoW is used to infer semantic meaning of documents. If we could translate an image into a text document we might be able to employ similar methods. A question arises: How does one make a text document from an image?

  \subsection{Processing Pipeline}	
    Tsai describes a well established pipeline that allows translation of images into text documents \cite{tsai2012bag}. It consists of the following steps: (1) region detection, (2) feature extraction, (3) vector quantization and (4) BoW histogram formation as can be seen in \refFig{fig:bow_pipeline}. The author summarises the most commonly used methods for performing these tasks.

@@

  \subsubsection{Region Detection}
    Characteristic region detection is the first step in any Bag of Words framework. Numerous detection methods have been developed, which might choosing the right one tricky. The most common detectors use a Harris corner detector or 1\textsuperscript{st} or 2\textsuperscript{nd} image derivatives. A Harris-Laplace detector is an example of a Harris-based detector. The Harris function is scale adapted and its outcome is subjected to a Laplacian-of-Gaussian (LoG) operator, which selects points in scale space. 2\textsuperscript{nd} image derivatives~---~Hessians can be combined with a LoG operator in order to further select points in the Hessian's determinant space.

    A number of more advanced recipes for salient region localisation have been developed and implemented. These include for example SIFT \cite{sift_keypoint}, SUSAN \cite{susan_keypoint} and ISS \cite{iss_keypoint}, with the vast majority implemented for the 2D domain and only a few generalised to 3D. A comparative study of detectors available in the PointCloud Library (PCL) can be found in \cite{pcl_keypoint_comparision, 3d_keypoint_eval}. All of the above, called sparse feature detectors, resort to selection of maxima in specific state spaces. An entirely different scheme is to use a dense feature detector, that is to take a uniformly sampled grid of points. Dense detectors are advantageous in that they sample slow changing regions in terms of image derivatives. Examples of such regions are a clear sky or a calm ocean. Li \emph{et al} showed that dense detectors generally outperform sparse ones \cite{fei2005bayesian}.

  \subsubsection{Feature Extraction}
    Suppose only keypoint's coordinates were given. The smallest affine transformation applied to the image would render the keypoints useless. Feature extraction is about describing a region in a way that is invariant to affine transformations, light intensity changes or colour saturation. Achieving all these properties simultaneously can be hard but methods that achieve some of them do exist. Keypoint description takes form of coordinates in a high-dimensional space. One of the most precise and repeatable algorithm is SIFT \cite{sift_features}, which is a 3D histogram of gradients structured as a 128-dimensional vector of floating point values. It is the most often extracted descriptor in BoW pipelines. Other methods include various colour descriptors, binary descriptors such as 512-dimensional GIST \cite{ponce2011cv}. There are techniques designed for 3D exclusively. Among them are Persistent Point Feature Histogram (PFH) \cite{pfh_rusu2008}, its faster alternative FPFH \cite{fpfh_rusu2009} and PFHRGB, which takes into account colour information, all implemented in PCL.

  \subsubsection{Vector Quantization}
    Vector quantization comprises of: (1)~codebook construction from a trainset and (2)~descriptor parsing, performed on a test set only on all data, depending on the setting. The simplest way of building a dictionary is to find patterns or regions within descriptors computed on a trainset and use their centroids or medoids as the codebook's entries. The parsing phase is about assigning a descriptor to one of the codebook elements (e.g. 1-nearest neighbour matching with a centroid). The parsed descriptor is called a \textit{visual word}. Finally, a histogram of visual words can be built. The histogram have a fixed size equal to the codebook size. Each bin of the histogram tells us how many visual words of a particular kind an image contains.

    The kMeans is the single most popular vector quantization algorithm used in the BoW pipelines \cite{tsai2012bag}. Developed in 1950's, it is well known and easy to implement. kMeans divides all data into a predefined amount of clusters and computes the their centroids. Many modifications and alternative versions emerged \cite{kmeans_jain2010data}. Some of them are: faster than the original \emph{approximate kmeans}, \emph{hierarchical kmeans}, which automatically chooses the final number of clusters and a \emph{soft kmeans} --- a variation that allows fuzzy alignment. Fuzzy alignment means that each point can belong to several clusters with different weights (e.g. proportional to the inverse square of a distance to a centroid).

    If computational cost is of no concern or if required precision is of the utmost priority, Gaussian Mixture Model (GMM) can be used. GMM partitions data into a set of clusters, finds their means and covariances. In the parsing step it computes a probability distribution of a descriptor over all the clusters. Fuzzy aligned histogram is built by taking the probability distribution as weights. GMM can be thought of as a generalisation of the kMeans algorithm. One drawback is massive computational cost in comparison with still expensive kMeans. Recently, Perronnin \textit{et al} proposed GMM based fisher-kernel vector quantization step yielding superior results \cite{fisher1, fisher2}.

    After choosing a vector quantization algorithm a codebook size have to be determined. With a standard kMeans implementation the dictionary size defines how much information is retained in a BoW histogram. A big trainset containing many classes with large intra-class and inter-class variance is likely to require a huge codebook. On the other hand, too big a dictionary might introduce quantization artefacts. One has to bear in mind that the vector quantization step is the most time-consuming part of the BoW pipeline. The highest computational complexity is associated with the codebook construction step and it is $O(n^3)$, where $n$ is the codebook size.

@@

  \subsection{Applications in Computer Vision}
    The Bag of Words approach is a tool than enables us to represent an image by a single feature vector of a chosen size. Its most popular applications in computer vision are Content Based Image Retrieval (CBIR) and Scene/Object Classification.

  \subsubsection{Content Based Image Retrieval}
    CBIR is a computer vision approach to image retrieval from large image databases. Tangelder \emph{et al} provides an overview of techniques applicable to 3D objects \cite{tangelder2008survey}. The task of image retrieval is to find a database entry that fulfil certain conditions. If we wanted to perform the task efficiently, we would have to meet the following criteria \cite{toldo2009bag}: (1) All entries should be indexed in a concise way, (2) a (dis)similarity measure should be provided and (3) an efficient search algorithm should be available.

    Database entries (images) should be indexed beforehand. Otherwise we would have to compare all the entries explicitly. An index must be compact as well as discriminative. Suitable algorithms can be divided into three general categories: (1) feature based methods, (2) graph based methods and (3) other methods.

    Feature based methods can be further divided into global features and local features. The global features take form of a single vector (or a point in a \emph{d} dimensional space). They are usually associated with object's mass or volume or distributions of these. Being easy to compute and straightforward to implement, their discriminative power is rather low --- they cannot be used for partial matching, but are well suited for an early preprocessing.

    Local features, on the other hand, describe object's characteristic regions. Their shape is similar to that of the global features, but instead of a single point in space there are multiple ones --- one for each considered region. Tangelder argues that local features based approaches are inefficient and lead to a complex indexing problem \cite{tangelder2008survey}. At the same time Toldo \emph{et al} and Li \emph{et al} show that Bag of Words local feature based approach has no such drawbacks. On the contrary --- it is easy to implement, efficient and provides state-of-the-art results.

    Feature based methods can be either local or global. Local features describe characteristic regions of an object, which have to be identified first. We have to carry out similar computation for each salient region. An outcome is a set of vectors, one vector for each region in \emph{d} dimensional space, where \emph{d} is dimensionality of the descriptor. Global features are generally easier (faster) to compute and more compact --- they take form of a single \emph{d} dimensional vector. Often global features are more computationally efficient and can be easily used for object indexing . Before the advent of BoW it was not clear how to use local features for indexing, nor there was any straight-forward dissimilarity measure available \cite{tangelder2008survey}. The Bag of Words approach solved these issues. What is more, it is easy to implement, efficient and provides state-of-the-art results \cite{li2010investigating}.

  \subsubsection{Scene and Object Classification}
    Scene and object classification are among the most popular issues of computer vision nowadays. We would like to perform those tasks automatically for the sake of surveillance, navigation or automatic image tagging. The very first attempt to use BoW in scene classification was made by Csurka \textit{et al} in 2004 \cite{csurka2004visual}. They were inspired by an idea of a ``texton'', a building block of texture introduced earlier in pattern recognition and texture classification. The authors suggested a BoW pipeline described above. Then, they fed the resulting BoW histograms into two classifiers: Support Vector Machine (SVM) and Na\`ve Bayes (both discussed below). Fei-Fei \emph{et al} refined the original approach by examining several keypoint detectors and descriptors \cite{fei2005bayesian}. Moreover, they developed a novel probabilistic graphical model for classification. They achieved accuracy of 76\% on a large 13 category dataset.

    One of the best entries in an ILSVRC2013\footnote{\url{http://www.image-net.org/challenges/LSVRC/2013/results.php}} challenge, attempting to classify millions of images into 1000 categories, was submitted by a fisher vector based BoW approach as well. The top-5\footnote{top-5 --- the classification result is successful if the ground-truth category is in one of the five most probable categories predicted} accuracy rate was close to 85\%.

@@

    To classify means to, given a set of categories, produce a category label for a given set of features \cite{ponce2011cv}. Image classification fits this description perfectly. The only difficulty lies in complexity of a raw image. Suppose we have a $320$ by $240$ three channel colour image. This amounts to a total of $230400$ dimensions, far too many to feed into any classifier directly. Fortunately, a Bag of Words histogram is a compact and discriminative intermediate representation that solves this issue. It can be used with any classifier. Classifiers can be two-class classifiers (e.g. binary) or multi-class classifiers. The latter are often a combination of several binary classifiers. The most popular classifiers are: k-Nearest Neighbours, Logistic Regression, Softmax Regression, Na\`ive Bayes and Support Vector Machine.

\section{ DATASETS }

  Datasets are essential for the following reasons: (1) Supervised learning algorithms have ravenous appetite for labeled data. An open database of labeled RGBD images would be of a great help in evaluation of algorithms. (2) If any comparison of algorithms is to be meaningful it should be performed on the same data. There is a multitude of RGB datasets available: ImageNet\footnote{\url{http://www.image-net.org/}}, PascalVoc\footnote{\url{http://pascallin.ecs.soton.ac.uk/challenges/VOC/}}, LabelMe\footnote{\url{http://labelme.csail.mit.edu/Release3.0/browserTools/php/dataset.php}} and SUN\footnote{\url{http://groups.csail.mit.edu/vision/SUN/}}, among others. Unfortunately, they do not contain any depth information and cannot be used for the purpose of object classification based on RGBD data. On the other hand, the vast majority of RGBD datasets are focused on tracking or instance-level recognition, have insufficient number of examples per category or have too few categories. Author managed to find only two datasets suited to his needs. They are: the Berkeley 3D Object (B3DO) dataset \cite{B3DO} and a dataset compiled by Zhang \emph{et al} at the University of Tokyo\cite{zhangcategory}. The latter will be abbreviated Tokyo dataset from now on.

@@

    The Berkeley 3D Object dataset, shown in \refFig{fig:b3do_objects}, was specifically designed for the purpose of object detection and classification. It consists of cluttered images in indoor environment, with many labelled and random objects per image. The images are often poorly lit and objects are partially occluded. There are around 50 classes with more than 20 examples in each of them. RGBD data are provided as pairs of colour images (8 bit RGB jpeg files) and depth maps (16 bit png files). Images are densely labelled --- for every pair of images there is an xml file with annotations. Each annotation states a category of an object as well as it's bounding box location. Since neither image segmentation nor object detection is addressed in this paper we had to extract objects from the images before the dataset could be used.

@@

    The Tokyo dataset, shown in \refFig{fig:tokyo}, is comprised of high quality images of single labelled objects in different settings, with changing viewpoint, object orientation and texture. Additional unlabelled object can be present, however, and should not be taken into account. No bounding boxes are given, which renders the classification task similar to the scene classification problem. The dataset is presented as colour images (8 bit RGB jpeg files) and depth maps (csv files with XYZ coordinates of each pixel). This dataset is aimed at the task of object classification in casual images.

\section{ EXPERIMENTS }
  There is a plenty of algorithms suitable for any of the four steps of Bag of Words image processing. Most of them have multiple tunable parameters and their performance might be interdependent on each other. Comprehensive evaluation is inevitable if an optimal combination is to be selected. Considering the PointCloud Library\footnote{\url{http://pointclouds.org/}} (PCL) alone yields far too many options to test, with 12 keypoint detectors and more than 20 feature descriptors. A literature review helped to narrow the scope of evaluation considerably. All experiments were run on a notebook with Intel Core i7 quad core 2.2 GHz CPU, 8 GB of RAM and a nVidia M540 GPU.

@@

  \subsection{Codebook}
    KMeans is used in almost every implementation of the Bag of Words image processing pipeline \cite{tsai2012bag, toldo2009bag}. Csurka \emph{et al} showed that the number of visual words has a significant impact on the final results \cite{csurka2004visual}. A number of tests were run to discover an optimal codebook size for the current setting.

@@

    The results from \refFig{fig:cluster_b3do} partially confirm Csurka's findings. The performance raises with the increasing number of centroids up to the point of 1500 centroids and 65.22\%. The discrepancies might be caused by the following factors: (1) ISS detector finds too many or irrelevant points, thus introducing noise or (2) The FPFH descriptor has too few dimensions (33) to be divided into more than 1500 regions in a meaningful way.

@@

    Everyone of the 343 images from the Tokyo dataset were used. Data were split into test and train set in proportions 1:2 in order to provide more training examples due to very limited numbers of objects in some classes. Even tough the overall accuracy achieved on the Tokyo dataset is similar in value to that of the B3DO's, the structure of the result is very different as shown in table \ref{tab:tokyo_conf_matrix}. It is clearly visible that there is a strong correlation between a per class accuracy and the number of entries in this class. The highest performance in the bicycle category is coupled with the largest number of entries. On the other end of the scale there are the printer, scanner and cart categories with 2, 2 and 4 entries respectively and 0\% classification accuracy. On top of that, the differences between some classes are marginal. The majority of carts and baskets were put into the bucket category. It does not surprise, for they are simply akin as can be seen in \refFig{fig:tokyo}.

\section{ CONCLUSION }

>>>>

\documentclass[12pt]{article}
\usepackage[hidelinks]{hyperref}
\usepackage{array}
\usepackage{booktabs}
\usepackage{cite}
\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subscript}[1]{\ensuremath{_{\textrm{#1}}}}
\usepackage{graphicx}
\usepackage{mathptmx}

@@

\title{3D Object Classification Based on RGBD Images}
\author{Adam Kosiorek}
\maketitle
Affiliation: Institute of Automation and Robotics, The Faculty of Mechatronics 
of Warsaw University of Technology \\
website: http://iair.mchtr.pw.edu.pl \\
Supervisor: Prof. dr hab. Barbara Siemiątkowska \\
Defense date: 11.02.2014

@@

 This paper introduces a Bag of Words (BoW) semantic object classification 
framework based on RGBD data registered by a Microsoft Kinect camera. We 
introduce a BoW image processing 
pipeline, describe it's steps in detail (region detection, feature 
extraction and vector quantization) and we outline the most successful 
algorithms. Algorithms with high performance potential were identified and 
evaluated. 
Salient region detection is done by SIFT, ISS3D and Uniform Sampling. 
Description is performed by FPFH, PFH and PFHRGB. Vector quantization was 
carried out by KMeans. An SVM with the RBF kernel trained with an 
All-vs-All method was used as a classifier. We achieved the accuracy of 65.22\% 
on 8 
categories of the Berkely 3D Object Dataset and 62.3\% on a 10 category dataset 
compiled by Zhang.

}

\section{ Introduction }

  It has never been easier to capture visual information. Popular storage 
services grow lighting fast due to terabytes of photographs and movies uploaded 
every day. The growth is so fast that hand-tagging and description, the 
traditional means of annotation, cease to suffice. They are ambiguous, 
emotional and rarely optimal. With no better solution at hand databases are 
becoming increasingly harder to browse. Similar issues occur in mobile 
robotics, where a robot has to understand it's enviornment if it is to interact 
with it. Both problems can be cast as a scene or object classification problem, 
one of the most popular computer vision issues in the last decade.

  This paper reviews the Bag of Words (or BoW for short) approach to image 
classification, evaluates available point cloud databases in terms of usability 
in scene classification studies and applies BoW model in the 3D domain. The 
rest of this work is organised as follows: Section 2 describes the Bag of Words 
technique in detail. Construction of an image model is discussed and the most 
popular algorithms are outlined and the problem of classification is 
introduced. Section 3 describes two renown RGBD image datasets suitable for 
evaluation of RGBD data-based object classifiers. Section 4 addresses the 
topics of experimental setup, conducted experiments and provides the results. 
Finally, Section 5 concludes this work.

\section{ Bag of Words }

  The Bag of Words model originated from the natural language processing domain 
\cite{tsai2012bag}. It 
assumes that word occurrence in a text document carries semantic information. 
It is an intermediate representation (in a form of a hisogram)
used in natural language processing, information retrieval and general data 
mining.

@@

  Figure \ref{fig:bow_example} shows a piece of text and its BoW model, with 
articles and punctuation marks left out. To compare different 
documents we need a global dictionary (or codebook), usually constructed by 
taking all unique words from available documents (a training set). If 
a dictionary is of any considerable size the resulting representation of 
especially small documents can be very sparse.

  In natural language processing BoW is used to infer semantic meaning of 
documents. If we could translate an image into a text document we might be able 
to employ similar methods. How do we do that?

  \subsection{From Images to Text Documents}	
The original pipeline consists of the following 
steps: (1) region detection, (2) feature extraction, (3) vector quantization 
and (4) BoW histogram formation as in figure \ref{fig:bow_pipeline}. Tsai 
summarises the most common methods used recently \cite{tsai2012bag}.

@@

  \subsubsection{Region Detection}
    Characteristic region detection is the first step in any Bag of Words 
framework. Numerous detection methods have been developed, which makes choosing 
the right one tricky. The most common detectors use a Harris corner detector or 
1\textsuperscript{st} or 2\textsuperscript{nd} image derivatives. A 
Harris-Laplace detector is an example of a Harris-based detector. The Harris 
function is scale adapted and its outcome is subjected to a 
Laplacian-of-Gaussian (LoG) operator, which selects points in scale space. 
2\textsuperscript{nd} image derivatives~---~Hessians can be combined with a LoG 
operator in order to further select points in the Hessian's determinant space.

    A number of more advanced recipes for salient region localisation have been 
developed and implemented. These include for example SIFT
and ISS, with the vast majority 
implemented for the 2D domain and only a few generalised to 3D. A comparative 
study of detectors available in the PointCloud Library (PCL) can be found in 
\cite{pcl_keypoint_comparision, 3d_keypoint_eval}. All of the above, called 
sparse feature detectors, resort to selection of maxima in specific state 
spaces. An entirely different scheme is to use a dense feature detector, that 
is to take a uniformly sampled grid of points. Dense detectors are advantageous 
in that they sample slow changing regions in terms of image derivatives. 
Examples of such regions are a clear sky or a calm ocean. It was
showed that dense detectors generally outperform sparse ones 
\cite{tsai2012bag}.

  \subsubsection{Feature Extraction}
    Suppose only keypoint's coordinates were given. The smallest affine 
transformation applied to the image would render the keypoints useless. Feature 
extraction is about describing a region in a way that is invariant to affine 
transformations, light intensity changes or colour saturation. Achieving all 
these properties simultaneously can be hard but methods that achieve some of 
them do exist. Keypoint description takes form of coordinates in a 
high-dimensional space. One of the most precise and repeatable algorithm is 
SIFT \cite{sift_features}, which is a 3D histogram of gradients structured as a 
128-dimensional vector of floating point values. It is the most often extracted 
descriptor in BoW pipelines. Other methods include various colour descriptors, 
binary descriptors such as 512-dimensional GIST. There are 
techniques designed for 3D exclusively. Among them are Persistent Point Feature 
Histogram (PFH) , its faster alternative FPFH 
 and PFHRGB, which takes into account colour information, 
all implemented in PCL.

  \subsubsection{Vector Quantization}
    Vector quantization comprises of: (1)~codebook construction and 
(2)~descriptor parsing.

To build a dictionary we have to describe the training set (the previous steps) 
and find patterns in the descriptor space. A codebook can then contain 
centroids (means or medoids) of these patterns. The parsing phase is about 
assigning a descriptor to one of the codebook elements (e.g. 1-nearest 
neighbour matching with a centroid). A parsed descriptor is called a 
\textit{visual word}. Finally, we build a histogram by counting how many times 
each visual word occured for an image. The result is a fixed-size histogram, 
whose each bin tells us how many times a visual word occured.

    The kMeans is the single most popular vector quantization algorithm used in 
the BoW pipelines \cite{tsai2012bag}. Developed in 1950's, it is well known and 
easy to implement. kMeans divides all data into a predefined amount of clusters 
and computes their centroids. Many modifications and alternative versions 
emerged \cite{kmeans_jain2010data}. Some of them are: faster than the original 
\emph{approximate kmeans}, \emph{hierarchical kmeans}, which automatically 
chooses the final number of clusters and a \emph{soft kmeans} --- a variation 
that allows fuzzy alignment. Fuzzy alignment means that each point can belong 
to several clusters with different weights (e.g. proportional to the inverse 
square of a distance to a centroid).

    If computational cost is of no concern or if required precision is of the 
utmost priority we can use Gaussian Mixture Model (GMM). It partitions data 
into a set of clusters, finds their means and covariances. In the parsing step 
it computes a probability distribution of a descriptor over all the clusters. 
Fuzzy aligned histogram is built by taking the probability distribution as 
weights. GMM can be thought of as a generalisation of the kMeans algorithm. One 
drawback is massive computational cost in comparison with still expensive 
kMeans. Recently, Perronnin \textit{et al} proposed GMM based fisher-kernel 
vector quantization step yielding superior results \cite{fisher1, fisher2}.

    After choosing a vector quantization algorithm a codebook size have to be 
determined. With a standard kMeans implementation the dictionary size defines 
how much information is retained in a BoW histogram. A big trainset containing 
many classes with large intra- and inter-class variance is likely to 
require a huge codebook. On the other hand, too big a dictionary might 
introduce quantization artefacts. We have to bear in mind that the vector 
quantization step is the most time-consuming part of the BoW pipeline. The 
highest computational complexity is associated with the codebook construction 
step and it is $O(n^3)$, where $n$ is the codebook size.

@@

@@

    To classify means to, given a set of categories, produce a category label 
for a given set of features. Image classification fits this 
description perfectly. The only difficulty lies in complexity of a raw image. 
Suppose we have a $320$ by $240$ three channel colour image. This amounts to a 
total of $230400$ dimensions, far too many to feed into any classifier 
directly. Fortunately, a Bag of Words histogram is a compact and discriminative 
intermediate representation that solves this issue. It can be used with any 
classifier. Classifiers can be two-class classifiers (e.g. binary) or 
multi-class classifiers. The latter are often a combination of several binary 
classifiers. The most popular classifiers are: k-Nearest Neighbours, Logistic 
Regression, Softmax Regression, Na\`ive Bayes and Support Vector Machine.

\section{ Datasets}

  Datasets are essential for the following reasons: (1) Supervised learning 
algorithms have ravenous appetite for labeled data. An open database of labeled 
RGBD images would be of great help in evaluation of algorithms. (2) If any 
comparison of algorithms is to be meaningful it should be performed on the same 
data. There is a multitude of RGB datasets available: 
ImageNet\footnote{\url{http://www.image-net.org/}}, 
PascalVoc\footnote{\url{http://pascallin.ecs.soton.ac.uk/challenges/VOC/}}, 
LabelMe\footnote{\url{
http://labelme.csail.mit.edu/Release3.0/browserTools/php/dataset.php}} and 
SUN\footnote{\url{http://groups.csail.mit.edu/vision/SUN/}}, among others. 
Unfortunately, they do not contain any depth information and cannot be used for 
the purpose of object classification based on RGBD data. On the other hand, the 
vast majority of RGBD datasets are focused on tracking or instance-level 
recognition, have insufficient number of examples per category or have too few 
categories. Author managed to find only two datasets suited to his needs. They 
are: the Berkeley 3D Object (B3DO) dataset \cite{B3DO} and a dataset compiled 
by Zhang \emph{et al} at the University of Tokyo\cite{zhangcategory}. The 
latter will be abbreviated Tokyo dataset from now on.

@@

    The Berkeley 3D Object dataset, shown in figure \ref{fig:b3do_objects}, was 
specifically designed for the purpose of object detection and classification. 
It consists of cluttered images in indoor environment, with many labelled and 
random objects per image. The images are often poorly lit and objects are 
partially occluded. There are around 50 classes with more than 20 examples in 
each of them. RGBD data are provided as pairs of colour images (8 bit RGB jpeg 
files) and depth maps (16 bit png files). Images are densely labelled --- for 
every pair of images there is an xml file with annotations. Each annotation 
states a category of an object as well as it's bounding box location. Since 
neither image segmentation nor object detection is addressed in this paper we 
had to extract objects from the images before the dataset could be used.

@@

    The Tokyo dataset, shown in figure \ref{fig:tokyo}, is comprised of high 
quality images of single labelled objects in different settings, with changing 
viewpoint, object orientation and texture. Additional unlabelled objects can be 
present, however, and should not be taken into account. No bounding boxes are 
given, which renders the classification task similar to the scene 
classification problem. The dataset is presented as colour images (8 bit RGB 
jpeg files) and depth maps (csv files with XYZ coordinates of each pixel). This 
dataset is aimed at the task of object classification in casual images.

\section{ Experiments }
  There is a plenty of algorithms suitable for any of the four steps of Bag of 
Words image processing. Most of them have multiple tunable parameters and their 
performance might be interdependent on each other. Comprehensive evaluation is 
inevitable if an optimal combination is to be selected. Considering the 
PointCloud Library\footnote{\url{http://pointclouds.org/}} (PCL) alone yields 
far too many options to test, with 12 keypoint detectors and more than 20 
feature descriptors. A literature review helped to narrow the scope of 
evaluation considerably. All experiments were run on a notebook with Intel Core 
i7 quad core 2.2 GHz CPU, 8 GB of RAM and a nVidia M540 GPU.

@@

  \subsection{Codebook}
    KMeans is used in almost every implementation of the Bag of Words image 
processing pipeline \cite{tsai2012bag}. The number of visual words can have a 
significant impact on the final results. A number of tests were run to discover 
an optimal codebook size for the current setting.

@@

    The results from figure \ref{fig:cluster_b3do} partially confirm Csurka's 
findings. The performance raises with the increasing number of centroids up to 
the point of 1500 centroids and 65.22\%. The discrepancies might be caused by 
the following factors: (1) ISS detector finds too many or irrelevant points, 
thus introducing noise or (2) The FPFH descriptor has too few dimensions (33) 
to be divided into more than 1500 regions in a meaningful way.

@@

    Everyone of the 343 images from the Tokyo dataset were used. Data were 
split into test and train set in proportions 1:2 in order to provide more 
training examples due to very limited numbers of objects in some classes. Even 
tough the overall accuracy achieved on the Tokyo dataset is similar in value to 
that of the B3DO's, the structure of the result is very different as shown in 
table \ref{tab:tokyo_conf_matrix}. It is clearly visible that there is a strong 
correlation between a per class accuracy and the number of entries in this 
class. The highest performance in the bicycle category is coupled with the 
largest number of entries. On the other end of the scale there are the printer, 
scanner and cart categories with 2, 2 and 4 entries respectively and 0\% 
classification accuracy. On top of that, the differences between some classes 
are marginal. The majority of carts and baskets were put into the bucket 
category. It does not surprise, for they are simply akin as can be seen in 
figure \ref{fig:tokyo}.

\section{ Conclusion }
